---
title: "Exploring gene clustering"
output: html_notebook
---


```{r}
suppressPackageStartupMessages({
  library(here)
  library(DESeq2)
  library(pheatmap)
  library(tidyverse)
  #library(RColorBrewer)
  library(vsn)
  library(genefilter)
  })
```

```{r}
source("utils.R")
source("fig_params.R")
#### Set a colour palette ####
#heat_colours <- brewer.pal(6, "YlOrRd")

plot_meanSd <- function(out_file, 
                        transformed_obj, 
                        w_in = width_in, 
                        h_in = height_in){
    # Open file
    svg(out_file, 
        width = w_in, 
        height = h_in)

    # Create the plot
    meanSdPlot(assay(transformed_obj))
    # Close the file
    dev.off()
}

#### Read in meta data ####
meta <- read_tsv(file = "experiment_metadata.tsv",
                 col_types = "ccc")

metadata <- create_meta(meta)

dds <- readRDS("DESeq_normalised.RDS")

#### Transform the data so that it's better visualised with heatmaps ####
rld <- rlog(dds, blind=TRUE)
vstd <- vst(dds, blind=TRUE)
# this gives log2(n + 1)
ntd <- normTransform(dds)




```

#### Store the transformed objects for further use ####
saveRDS(rld, file = snakemake@output[["rlog_transformed"]])
saveRDS(vstd, file = snakemake@output[["vs_transformed"]])
saveRDS(ntd, file = snakemake@output[["norm_transformed"]])

#### Plot meanSd ####
# see https://bioconductor.org/packages/release/bioc/vignettes/DESeq2/inst/doc/DESeq2.html#effects-of-transformations-on-the-variance
plot_meanSd(snakemake@output[["rlog_meanSdPlot"]], rld)
plot_meanSd(snakemake@output[["vs_meanSdPlot"]], vstd)
plot_meanSd(snakemake@output[["norm_meanSdPlot"]], ntd)

```{r}
anno <- as.data.frame(colData(vstd)[, c("condition")])
condition <- colData(vstd)$condition
anno2 <- as.data.frame(condition)
rownames(anno2) <- rownames(colData(vstd))
```



```{r}
create_annotation <- function(dESeqTransformObject){
  #anno <- as.data.frame(colData(dESeqTransformObject)[, c("condition")])
  condition <- colData(dESeqTransformObject)$condition
  anno <- as.data.frame(condition)
  rownames(anno) <- rownames(colData(dESeqTransformObject))
  anno
}

extract_n_highest_var_across_samples <- function(dESeqTransformObject, n){
  topVarGenes <- head(order(rowVars(assay(dESeqTransformObject)), decreasing = TRUE), n)
}

get_deviation_from_avg_across_samples <- function(dESeqTransformObject, pos_of_genes_of_interest){
  mat <-assay(dESeqTransformObject)[ pos_of_genes_of_interest, ]
  mat <- mat - rowMeans(mat)
}

#### Plot heatmap ####
plot_heatmap <- function(mat, 
                         output_dest,
                         annot = annotation,
                         heatmap_title = "Clustering of 20 genes with the variance across samples",
                         w_in = width_in,
                         h_in = height_in) {
  svg(output_dest, 
      width = w_in,
      height = h_in)
  
  # Plot heatmap
  corr_heatmap <- pheatmap(mat,
                         main = heatmap_title,
                         annotation = annot)
  dev.off()
}

plot_heatmap(mat = get_deviation_from_avg_across_samples(dESeqTransformObject = vstd,
                                                         pos_of_genes_of_interest = extract_n_highest_var_across_samples(dESeqTransformObject = vstd,
                                                                                                                         n = 20)),
             output_dest = "heatmap.svg",
             annot = create_annotation(dESeqTransformObject = vstd))

```

```{r}
colorRampPalette(rev(brewer.pal(n = 7, name = "RdYlBu")))(100)
```













#### Do the actual plotting ####
topVarGenes <- head(order(rowVars(assay(rld)), decreasing = TRUE), 20)





plot_heatmap(mat_cor = compute_pairwise_correlation(rld),
             output_dest = snakemake@output[["rlog_transformed_corr_heatmap"]],
             mdata = metadata)
plot_heatmap(mat_cor = compute_pairwise_correlation(vstd),
             output_dest = snakemake@output[["vs_transformed_corr_heatmap"]],
             mdata = metadata)